---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(tidyverse)
library(stringr)
library(lubridate)
```


```{r}
loan2 <- read.csv(file.choose(), header = T)
```


As the grade decreses, the interest rate increases 
```{r}
ggplot(loan2) + 
  geom_boxplot(aes(x = grade , y = int_rate , fill = grade)) + 
  labs(y = 'Interest Rate' , x = 'Grade')
```


Graph displays proportion of defaults and fully paid loans per grade
```{r}
ggplot(loan2) + 
  geom_bar(aes(x = grade , y = ..count.. , fill = factor(loan_outcome , c(1 , 0) , c('Default' , 'Fully Paid'))), position = "fill")
```

```{r}
colnames(loan2)
```

Create dataset with important variables
```{r}
loan2 <- loan2%>%
  select(loan_outcome , loan_amnt , int_rate , grade , emp_length , home_ownership , 
               annual_inc , term, issue_d)
```
```{r}
loan7 <- loan2%>%
  select(loan_outcome , loan_amnt , int_rate , grade , emp_length , home_ownership , 
               annual_inc , term, issue_d, last_pymnt_d)
```

Lets see default rate by income...

Startify income by creating a new variable
```{r}
#use quartiles to find the breaks
summary(loan2$annual_inc)
loan2 <- loan2%>%
  mutate(income_category = with(loan2, cut(annual_inc, breaks = c(0, 45000, 65000, 90000, 10999200), labels = c('45000 or less', '45001-65000',
  '65001-90000', '90000+'))))
```
Default rate increases slightly as income decreses
```{r}
ggplot(loan2) + 
  geom_bar(aes(x = income_category , y = ..count.. , fill = factor(loan_outcome , c(1 , 0) , c('Default' , 'Fully Paid'))), position = "fill")+
  coord_flip()
```

length of employment does not seem to affect default rate
```{r}
loan2$emp_length <- as.factor(loan2$emp_length)
ggplot(loan2) + 
  geom_bar(aes(x = emp_length , y = ..count.. , fill = factor(loan_outcome , c(1 , 0) , c('Default' , 'Fully Paid'))), position = "fill")+
  coord_flip()
```
Appears that home_ownership does not have too large an effect on default rate
```{r}
loan2$home_ownership <- as.factor(loan2$home_ownership)
ggplot(loan2) + 
  geom_bar(aes(x = home_ownership , y = ..count.. , fill = factor(loan_outcome , c(1 , 0) , c('Default' , 'Fully Paid'))), position = "fill")+
  coord_flip()
```

```{r}
#install.packages("plyr")
library(plyr)
```

20.07% have defaulted
```{r}
count(loan2, 'loan_outcome')
261686/ (1041952+261686)
```


Unlike the bar graph above, these fractions show us that there actually does seem to be a relationship between income and default rate 
```{r}
loan2%>%
  group_by(loan_outcome, income_category)%>%
  dplyr::summarise(count=n())
#45000 or less
75634/(246981+75634)	
#45001-65000
75353/(279181+75353)	
#65001-90000
58124/ (244947 + 58124)
#90000+
52515	/ (270601	+ 52515)
```



```{r}
loan2
```


60 months twice as likely to defualt than 36 months 
```{r}
loan%>%
  group_by(term, loan_outcome)%>%
  dplyr::summarize(count = n())
#36 months 
159169/ (829605+159169)
#60 months
102517 / (212347+102517)
```

Default rate per emp_length level, emp_length does not seem to have that big an impact...hb during a recession?
```{r}
loan2
loan2%>%
  group_by(emp_length, loan_outcome)%>%
  dplyr::summarize(count = n())

#<1
21622 / (82930+21622) = .2
#1
17692/(67986+17692)
#2
23482/(94343+23482)
#3
20937/ (20937+83267)
#10+
80933/ (80933+347620)
```


```{r}
loan3 <- loan2%>%
  mutate(issue_dc= issue_d)
loan3$issue_d <- paste(loan3$issue_d, "-01", sep = "")
loan3
```

```{r}

loan7$issue_d <- paste(loan7$issue_d, "-01", sep = "")
loan7
```
```{r}
loan7$last_pymnt_d <- paste(loan7$last_pymnt_d, "-01", sep = "")
loan7
```


```{r}
loan3$issue_d <- str_replace(loan3$issue_d, "Dec", "12")
loan3$issue_d <- str_replace(loan3$issue_d, "Jan", "01")
loan3$issue_d <- str_replace(loan3$issue_d, "Feb", "02")
loan3$issue_d <- str_replace(loan3$issue_d, "Mar", "03")
loan3$issue_d <- str_replace(loan3$issue_d, "Apr", "04")
loan3$issue_d <- str_replace(loan3$issue_d, "May", "05")
loan3$issue_d <- str_replace(loan3$issue_d, "Jun", "06")
loan3$issue_d <- str_replace(loan3$issue_d, "Jul", "07")
loan3$issue_d <- str_replace(loan3$issue_d, "Aug", "08")
loan3$issue_d <- str_replace(loan3$issue_d, "Sep", "09")
loan3$issue_d <- str_replace(loan3$issue_d, "Oct", "10")
loan3$issue_d <- str_replace(loan3$issue_d, "Nov", "11")
```

```{r}
loan7$issue_d <- str_replace(loan7$issue_d, "Dec", "12")
loan7$issue_d <- str_replace(loan7$issue_d, "Jan", "01")
loan7$issue_d <- str_replace(loan7$issue_d, "Feb", "02")
loan7$issue_d <- str_replace(loan7$issue_d, "Mar", "03")
loan7$issue_d <- str_replace(loan7$issue_d, "Apr", "04")
loan7$issue_d <- str_replace(loan7$issue_d, "May", "05")
loan7$issue_d <- str_replace(loan7$issue_d, "Jun", "06")
loan7$issue_d <- str_replace(loan7$issue_d, "Jul", "07")
loan7$issue_d <- str_replace(loan7$issue_d, "Aug", "08")
loan7$issue_d <- str_replace(loan7$issue_d, "Sep", "09")
loan7$issue_d <- str_replace(loan7$issue_d, "Oct", "10")
loan7$issue_d <- str_replace(loan7$issue_d, "Nov", "11")
```



```{r}
loan3
loan3$issue_d <- myd(loan3$issue_d)

```
```{r}

loan7$issue_d <- myd(loan7$issue_d)
```


```{r}
loan3
```

```{r}
write.csv(loan7, "clean_loan2.csv")
```




```{r}
ggplot(loan3) + 
  geom_histogram(aes(x = issue_d , y = ..count.. , fill = factor(loan_outcome , c(1 , 0) , c('Default' , 'Fully Paid'))), bins = 30, position = "fill")
```
Default rate seems to get higher as the years progress
```{r}
cdplot(as.factor(loan_outcome) ~ issue_d, data=loan3, col=c("cornflowerblue", "orange"), main="Conditional density plot")
```
emp_length, loan_outcome, issue_d

```{r}
bond2 <- read_csv("2yrbond.csv")
```

```{r}
bond2
bond2$DATE <- mdy(bond2$DATE)
bond2
```





```{r}
bond2
```


```{r}
ggplot(bond2)+
  geom_line(aes(x = DATE, y = as.double(DGS2)))
```
```{r}

```


create a default rate variable for each money

```{r}
loan3
loan3%>%
  select(issue_dc, loan_outcome)%>%
  dplyr::group_by(issue_dc, as.factor(loan_outcome))%>%
  summarize(npl = n())
  
```

```{r}
#loan3

merge1 <- aggregate(loan3$loan_outcome, by=list(Date=loan3$issue_d), FUN=length)
merge2 <- aggregate(loan3$loan_outcome, by=list(Date=loan3$issue_d), FUN=sum)
merge2
```
```{r}
merge1
merge2
merge2 <- merge2%>%
  mutate(y=x)%>%
  select(Date, y)
```
```{r}
merge1
merge2
```
471 values, take value of previus day
```{r}
bond2
#length(bond2$DGS2[bond2$DGS2=="."])
bond2$DGS2 <- na_if(bond2$DGS2, ".")
bond2
```

```{r}
#library(zoo)
bond2$DGS2 <- na.locf(bond2$DGS2, fromLast = TRUE)
```

```{r}
complete_bond <- bond2
```


```{r}

loan4 <- full_join(merge1, merge2,by = "Date" )

loan4 <- loan4%>%
  mutate(total = x, num_default = y)%>%
  select(Date, total, num_default)
```

```{r}
loan4
#bond2%>%
 # mutate(day = day(DATE))%>%
  #filter(day==1)%>%
  #select(DGS2, issue_d)
bond2
#does not include a lot of jan 1 bonds FYI
bond2 <- bond2%>%
  mutate(Date = DATE) #year = year(Date), day = day(Date))%>%
  #select(Date, DGS2, year, day)%>%
  f#ilter(year >= 2007 ,day == 1)

```
```{r}
#bond2 <- bond2%>%
  #select(Date, DGS2)
bond2
loan4
```




```{r}

loan_bond <- full_join(loan4, bond2, by = "Date")
loan_bond$DGS2 <- na.locf(loan_bond$DGS2, fromLast = TRUE)

```
```{r}
loan_bond<- loan_bond%>%
  filter(!is.na(total))%>%
  select(Date, total, num_default, DGS2)
```


```{r}
loan_bond <- loan_bond%>%
  mutate(default_rate = num_default / total *100)

```

```{r}
loan_bond$DGS2 <- as.double(loan_bond$DGS2)
loan_bond
```

```{r}
ggplot(loan_bond, aes(x=Date))+
  geom_smooth(aes(y=default_rate), color = "blue")+
  geom_smooth(aes(y=DGS2))
```

```{r}
ggplot(loan_bond)+
  geom_smooth(aes(x=DGS2, y=default_rate))+
  geom_point(aes(x=DGS2, y=default_rate))
```
```{r}
cor(loan_bond$default_rate, loan_bond$DGS2)
```

```{r}
temp <- loan_bond%>%
  filter(year(Date) <=2008)
temp
cor(temp$default_rate, temp$DGS2)
ggplot(temp)+
  geom_point(aes(x= DGS2, y = default_rate))
```
```{r}
write.csv(loan_bond, "loan_bond.csv")
```
```{r}
write.csv(complete_bond, "complete_bond.csv")
```
```{r}
write.csv(loan3, "clean_loan.csv")

```

lets see how bond yield affects default based on income....