---
title: "Model Comparison"
output: html_notebook
---


Logistic Regression with Elastic Net Regularization
```{r}
#model #model
#res #train
roc_lr <- roc((loan_outcome) ~ as.numeric(lr_res), data = test1)
plot(roc_lr, col = 'purple')
roc_lr


```


Linear Discriminant Analysis
```{r}
#lda #model
roc_lda <- roc(as.factor(loan_outcome) ~ as.numeric(lda_res), data = test1)
plot(roc_lda, col = "purple")
roc_lda
```


Cost-sensitive Gradient Boosting Machine
interaction debth = 7 
shrinkage = 0.01
weights = (1,10)
```{r}
#gbm.tuned #model
#gbm_res <- gbm_test_tuned #test
roc_gbm <- roc(as.factor(loan_outcome) ~ gbm_res, data = test1)
plot(roc_gbm, col = "purple")
roc_gbm
```


Random Forest, 375 Trees
```{r}
#rf_tune #model
#rf_res #test
roc_rf <- roc(as.factor(loan_outcome) ~ as.numeric(rf_res), data = test1)
plot(roc_rf)
roc_rf
```

Ensemble Classifier
```{r}
#predict_loan_status_ensemble #test
roc_ensemble <- roc(as.factor(loan_outcome) ~ as.numeric(ensemble_res), data = test1)
plot(roc_ensemble)
roc_ensemble
```

Compare ROCs
```{r}
plot(roc_lr,legacy.axes = TRUE,col="yellow",main="ROC compare")
plot(roc_lda,legacy.axes = TRUE,col="purple",add=TRUE)
plot(roc_rf,legacy.axes = TRUE,col="green",add=TRUE)
plot(roc_gbm,legacy.axes = TRUE,col="pink",add=TRUE)
plot(roc_ensemble,legacy.axes = TRUE,col="blue",add=TRUE)



legend("bottomright",legend=c("LR","LDA","RF","GBM","ensemble"),fill=c("yellow","purple","green","pink","blue"))

```




LC grade percentage 
```{r}
LC <- compare%>%group_by(grade)%>%summarize(percent = n()/6387, precision =1- mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd), poverty = mean(poverty_rate), growth = mean(growth), hs = mean(hs), bach = mean(bach), inc = median(annual_inc))
compare%>%filter(loan_outcome==1&grade=="A")%>%summarize(grade = "A", lgd = 1-mean(actual_return))
compare%>%filter(loan_outcome==1&grade=="B")%>%summarize(grade = "B", lgd = 1-mean(actual_return))
compare%>%filter(loan_outcome==1&grade=="C")%>%summarize(grade = "C", lgd = 1-mean(actual_return))
compare%>%filter(loan_outcome==1&grade=="D")%>%summarize(grade = "D", lgd = 1-mean(actual_return))
compare%>%filter(loan_outcome==1&grade=="E")%>%summarize(grade = "E", lgd = 1-mean(actual_return))
compare%>%filter(loan_outcome==1&grade=="F")%>%summarize(grade = "F", lgd = 1-mean(actual_return))
compare%>%filter(loan_outcome==1&grade=="G")%>%summarize(grade = "GR", lgd = 1-mean(actual_return))

LC
compare%>%summarize(int = median(int_rate))
```

Create ensemble table
```{r}
ce <- cbind(compare,ensemble_res)
e_a<- ce%>%filter(ensemble_res< .13)%>%summarize(grade = "A", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd), poverty = mean(poverty_rate), growth = mean(growth), hs = mean(hs), bach = mean(bach), inc = median(annual_inc))
ce%>%filter(loan_outcome==1&ensemble_res<.214)%>%summarize(lgd = 1-mean(actual_return))

e_b<- ce%>%filter( ensemble_res >= .13 & ensemble_res < .21)%>%summarize(grade = "B", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd), poverty = mean(poverty_rate), growth = mean(growth), hs = mean(hs), bach = mean(bach), inc = median(annual_inc))
ce%>%filter(loan_outcome==1&ensemble_res >= .13 & ensemble_res < .21)%>%summarize(lgd = 1-mean(actual_return))

e_c<- ce%>%filter( ensemble_res >= .21 & ensemble_res < .27)%>%summarize(grade = "C", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd), poverty = mean(poverty_rate), growth = mean(growth), hs = mean(hs), bach = mean(bach), inc = median(annual_inc))
ce%>%filter(loan_outcome==1& ensemble_res >= .21 & ensemble_res < .27)%>%summarize(lgd = 1-mean(actual_return))

e_d<- ce%>%filter( ensemble_res >= .27 & ensemble_res < .34)%>%summarize(grade = "D", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd), poverty = mean(poverty_rate), growth = mean(growth), hs = mean(hs), bach = mean(bach), inc = median(annual_inc))
ce%>%filter(loan_outcome==1&  ensemble_res >= .27 & ensemble_res < .34)%>%summarize(lgd = 1-mean(actual_return))

e_e<- ce%>%filter( ensemble_res >= .34 & ensemble_res < .41)%>%summarize(grade = "E", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd), poverty = mean(poverty_rate), growth = mean(growth), hs = mean(hs), bach = mean(bach), inc = median(annual_inc))
ce%>%filter(loan_outcome==1&  ensemble_res >= .34 & ensemble_res < .41)%>%summarize(lgd = 1-mean(actual_return))

e_f<- ce%>%filter( ensemble_res >= .41 & ensemble_res < .43)%>%summarize(grade = "F", percent = n()/6387, precision= 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd),poverty = mean(poverty_rate), growth = mean(growth), hs = mean(hs), bach = mean(bach), inc = median(annual_inc))
ce%>%filter(loan_outcome==1&  ensemble_res >= .41 & ensemble_res < .43)%>%summarize(lgd = 1-mean(actual_return))
e_g<- ce%>%filter( ensemble_res >= .43)%>%summarize(grade = "G", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd),poverty = mean(poverty_rate), growth = mean(growth), hs = mean(hs), bach = mean(bach), inc = median(annual_inc))

ce%>%filter(loan_outcome==1&  ensemble_res >= .43)%>%summarize(lgd = 1-mean(actual_return))
E <- rbind(e_a, e_b, e_c, e_d, e_e, e_f, e_g)
E
```


Create RF table 
```{r}
cr <- cbind(compare, rf_res)
rf_a<- cr%>%filter( rf_res < .125)%>%summarize(grade = "A", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
rf_a
rf_b<- cr%>%filter( rf_res >= .125 & rf_res < .24)%>%summarize(grade = "B", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
rf_b
rf_c<- cr%>%filter( rf_res >= .24 & rf_res < .35)%>%summarize(grade = "C", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
rf_c
rf_d<- cr%>%filter( rf_res >= .35 & rf_res < .38)%>%summarize(grade = "D", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
rf_d
rf_e<- cr%>%filter( rf_res >= .38& rf_res < .46)%>%summarize(grade = "E", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
rf_e
rf_f<- cr%>%filter( rf_res >= .46& rf_res < .50)%>%summarize(grade = "F", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
rf_f
rf_g<- cr%>%filter( rf_res >= .50)%>%summarize(grade = "G", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
rf_g
RF <- rbind(rf_a, rf_b, rf_c, rf_d, rf_e, rf_f, rf_g)
RF
```


Create LDA table
```{r}
cl <- cbind(compare, lda_res)
lda_a<- cl%>%filter(lda_res<.089)%>%summarize(grade = "A", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lda_a
lda_b<- cl%>%filter(lda_res>=.089 & lda_res < .19)%>%summarize(grade = "B", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100,interest = mean(int_rate), lgd = mean(lgd))
lda_b
lda_c<- cl%>%filter(lda_res>=.19 & lda_res < .28)%>%summarize(grade = "C", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lda_c
lda_d<- cl%>%filter(lda_res>=.28 & lda_res < .34)%>%summarize(grade = "D", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lda_d
lda_e<- cl%>%filter(lda_res>=.34& lda_res < .50)%>%summarize(grade = "E", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lda_e
lda_f<- cl%>%filter(lda_res>=.50 & lda_res < .63)%>%summarize(grade = "F", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lda_f
lda_g<- cl%>%filter(lda_res>=.63)%>%summarize(grade = "G", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lda_g
LDA <- rbind(lda_a, lda_b, lda_c, lda_d, lda_e, lda_f, lda_g)
LDA
```




Create GBM table 
```{r}
cg <- cbind(compare, gbm_res)
gbm_a<- cg%>%filter(gbm_res<.112)%>%summarize(grade = "A", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
gbm_a
gbm_b<- cg%>%filter(gbm_res>=.112& gbm_res < .19)%>%summarize(grade = "B", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
gbm_b
gbm_c<- cg%>%filter(gbm_res>=.19& gbm_res < .24)%>%summarize(grade = "C", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
gbm_c
gbm_d<- cg%>%filter(gbm_res>=.24 & gbm_res < .37)%>%summarize(grade = "D", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
gbm_d
gbm_e<- cg%>%filter(gbm_res>=.37 & gbm_res < .45)%>%summarize(grade = "E", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
gbm_e
gbm_f<- cg%>%filter(gbm_res>=.45 & gbm_res < .50)%>%summarize(grade = "F", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
gbm_f
gbm_g<- cg%>%filter(gbm_res>=.50)%>%summarize(grade = "G", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100,interest = mean(int_rate), lgd = mean(lgd))
gbm_g
GBM <- rbind(gbm_a, gbm_b, gbm_c, gbm_d, gbm_e, gbm_f, gbm_g)
GBM
```


Create LR table
```{r} 
cl <- cbind(compare, lr_res)
lr_a <- cl%>%filter(lr_res<.128)%>%summarize(grade = "A", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lr_a
lr_b <- cl%>%filter(lr_res>=.128& lr_res <.21)%>%summarize(grade = "B", percent = n()/6387, precision = 1-mean(loan_outcome), return= (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lr_b
lr_c <- cl%>%filter(lr_res>=.21& lr_res <.26)%>%summarize(grade = "C", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lr_c
lr_d <- cl%>%filter(lr_res>=.26& lr_res <.28)%>%summarize(grade = "D", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lr_d
lr_e <- cl%>%filter(lr_res>=.29& lr_res <.39)%>%summarize(grade = "E", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lr_e
lr_f <- cl%>%filter(lr_res>=.39& lr_res <.45)%>%summarize(grade = "F", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lr_f
lr_g <- cl%>%filter(lr_res>=.45)%>%summarize(grade = "G", percent = n()/6387, precision = 1-mean(loan_outcome), return = (((sum(total_pymnt)+sum(recoveries))/sum(loan_amnt))-1)*100, interest = mean(int_rate), lgd = mean(lgd))
lr_g
LR <- rbind(lr_a, lr_b, lr_c, lr_d, lr_e, lr_f, lr_g)
LR
```


Merge Tables 
```{r}
LR$model = "lr"
GBM$model = "gbm"
LDA$model = "lda"
RF$model = "rf"
E$model = "ensemble"

LC$model = "lc"
all_models = rbind(LR, GBM, LDA, RF, E, LC)
```

Interest
```{r}
ggplot(all_models,aes(x=grade,y=interest,group=model)) + 
    geom_line(aes(color = model)) + 
    geom_point() +
    xlab("Grade") +
    ylab("Interest (%)") +
    labs(title = "Interest for each Grade Classification")
```

Return
```{r}
ggplot(all_models,aes(x=grade,y=percent,group=model)) + 
    geom_line(aes(color = model)) + 
    geom_point() +
    xlab("Grade") +
    ylab("Percent Suggested (%)") +
    labs(title = "Percent Suggested for each Grade Classification")
```


Percent Suggested
```{r}
ggplot(all_models,aes(x=grade,y=return,group=model)) + 
    geom_line(aes(color = model)) + 
    geom_point() +
    xlab("Grade") +
    ylab("Percent Suggested (%)") +
    labs(title = "Percent Suggested for each Grade Classification")
```




